---
title: "R Project 2"
author: "Sitong Chen"
date: "11/30/2021"
output: "html_document"
---
## Info550 - R project
####  Two dataset are used in this project: movies and ratings, I would like to explore the rating distribution by time and by the number of users, clean the movies dataset and see how many movies fall into each genre
```{r}
library(here)
here::i_am("R/Info550_project.Rmd")
knitr::opts_knit$set(root.dir = here::here())
```

```{r, message=FALSE}
library(readr)
library(tidyverse)
library(ggplot2)
library(data.table)
library(reshape2)
```

```{r,}
movies <- read.csv(here::here('data','movies.csv'),stringsAsFactors=FALSE)
ratings <- read.csv(here::here('data',"ratings.csv"))
str(movies)
```

```{r}
summary(movies)
head(movies)
```

```{r}
summary(ratings)
head(ratings)
```
#### Combine ratings and movies together by movieId and create some graphs
```{r}
movie_rating <- merge(ratings,movies,by="movieId")
```

#### Clean the date
```{r}
library(lubridate)
tibble('Initial Date' = date(as_datetime(min(movie_rating$timestamp),origin="1970-01-01")),
       'Final Date' = date(as_datetime(max(movie_rating$timestamp),origin="1970-01-01"))) %>%
  mutate(Period = duration(max(movie_rating$timestamp)-min(movie_rating$timestamp)))
```
#### Plot a bar graph to see the number of rating distribution from 1996 till 2016
```{r,message=FALSE}
library(ggthemes)
library(scales)
movie_rating %>% mutate(year = year(as_datetime(timestamp, origin="1970-01-01"))) %>%
  ggplot(aes(x=year)) +
    geom_histogram(color = "white") + 
    ggtitle("Rating Distribution Per Year") +
    xlab("Year") +
    ylab("Number of Ratings") +
    scale_y_continuous(labels = comma) + 
    theme_economist()
```

#### Plot a bar graph to see the rating distribution based on ratings from 1 to 5
```{r}
summary(movie_rating$rating)
```

```{r,message=FALSE}
# Count the number for each ratings
rating_dist <- movie_rating %>% group_by(rating) %>% summarize(n=n())
g <- ggplot(rating_dist, aes(x=rating, y=n))+geom_col()+labs(title="Ratings Distribution", y="Count",x="Ratings")
g
```

#### Investigate the users to see how many users rate over 1000 movies
##### 15 userid rate over 1000 movies
```{r}
movie_rating %>% group_by(userId) %>%
  summarise(n=n()) %>% 
  arrange(n) %>%
  filter(n>=1000)
```

#### Distribution of users with their rating count
```{r, message=FALSE}
movie_rating %>% group_by(userId)%>%
  summarise(n=n()) %>%
  ggplot(aes(n)) + 
  geom_histogram(color = "white") + 
  scale_x_log10()+
  ggtitle("Distribution of Users")+
  xlab("Number of Ratings") +
  ylab("Number of Users")+
  scale_y_continuous(labels=comma)+
  theme_economist()
```

#### Create a genre table with all the genres listed in movies dataset
##### I need to convert the genres present in the movies dataframe into a more usable format by the users.
##### To create a one-hot encoding to create a matrix that composes of corresponding genres for each of the films. 
```{r}
movies_genre <- as.data.frame(movies$genres, stringAsFactors = FALSE)
movies_genre2 <- as.data.frame(tstrsplit(movies_genre[,1],'[|]',type.convert = TRUE),stringAsFactors = FALSE)
colnames(movies_genre2) <- c(1:10)
genre_list <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")
genre_mat1 <- matrix(0,10330,18)
genre_mat1[1,]<-genre_list
colnames(genre_mat1) <- genre_list

for (index in 1:nrow(movies_genre2)){
  for (col in 1:ncol(movies_genre2)){
    genre_col = which(genre_mat1[1,]==movies_genre2[index,col])
    genre_mat1[index+1, genre_col] <- 1
  }
}

genre_mat2 <- as.data.frame(genre_mat1[-1,],stringAsFactors = FALSE) #remove the first row of genre list

for (col in 1:ncol(genre_mat2)){
  genre_mat2[,col] <- as.integer(genre_mat2[,col])
}

str(genre_mat2)
```


#### Create a dataframe specifying the genre present in our list so that we can see which genres this movie belongs to. 
```{r}
Movie_genre <- cbind(movies[,1:2],genre_mat2)
head(Movie_genre)
```
#### How many movies fall into each genre, which genre has the most movies and which genre has the least?
#### We can see from the output below, Drama has the most movies, and Film-Noir has the least. 
```{r}
Total_movie_genre <- as.data.frame(t(Movie_genre[,3:20]))
Total_movie_genre$genre <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")
Total_movie_genre <- Total_movie_genre %>%mutate(total_movies = rowSums(Total_movie_genre[1:10329]))%>%select(c(genre,total_movies))
Total_movie_genre %>% arrange(total_movies,descending=TRUE)

```

